### 2018-08-14 doc/issre2018/conclusions.tex
Prepared camera ready

\section{Conclusions}
\label{sec:concl}

In this paper we have introduced a novel search-based JS unit testing framework, called \emph{JEDI}. Notably, our framework is able to generate an arbitrary DOM inputs that are syntactically valid and compliant with the recent HTML5 specification. We present a test generation algorithm, called ``genetic with restart'', which is capable of escaping plateaus by concretizing the search target. Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation followed by the significance study, which has confirmed the effectiveness and efficiency of our framework in branch coverage testing.